Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-03-07 latex/MPP21/paper.tex
add GPT-3 comment
add example expression for bini

Our goal in this paper is to demonstrate that APA algorithms can offer practical performance improvements for applications that are tolerant to error in matrix multiplications, notably the training phase of neural networks.
Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.

@@

\GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}
We highlight several properties of the rule which are common across all APA algorithms we consider.
First, the rule requires fewer multiplications than the classical one (rank 10 instead of 12 for $\dims{3,2,2}$ here).
Next, each multiplication is between a linear combination of entries of $A$ and a linear combination of entries of $B$, each output entry is computed as a linear combination of the outputs of the multiplications, and each coefficient in the linear combinations is a (Laurent) polynomial in $\lambda$.
The coefficients include both positive and negative powers of $\lambda$, which explains why small values of $\lambda$ can lead to significant roundoff error.
For Bini's algorithm all coefficients are monomial with degree between $-1$ and $1$.

>>>>

Our goal in this paper is to demonstrate that APA algorithms can offer practical performance improvements for applications that are tolerant to error in matrix multiplications, notably the training phase of neural networks.
Training large neural networks is computationally expensive and the cost has spurred a surge of research into more efficient hardware, better algorithms, and techniques for trading off accuracy for performance.
For instance, low-precision arithmetic has been shown to decrease running time with little to no effect on the ultimate learning task \cite{GAGN15,HCSEB17}, and new floating point formats have been developed and supported in hardware to implement highly efficient low- and mixed-precision computation \cite{KM+19,YWC20}.
\JW{Open-AI's most recent language model, GPT-3, spent the equivalent of days training on a petaflop machine. Even a slight reduction in flops could save a large amount of time and reduce server costs.}\cite{BM+20}

@@

We can also expressed for $\hat{C}_{11}$ as $\hat C_{11} = A_{11} \cdot B_{11} + A_{12} \cdot B_{21} - \lambda (A_{12} \cdot B_{11})$, where $E = A_{12} \cdot B_{11}$ is the error matrix.
\GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}
We highlight several properties of the rule which are common across all APA algorithms we consider.
First, the rule requires fewer multiplications than the classical one (rank 10 instead of 12 for $\dims{3,2,2}$ here).
Next, each multiplication is between a linear combination of entries of $A$ and a linear combination of entries of $B$, each output entry is computed as a linear combination of the outputs of the multiplications, and each coefficient in the linear combinations is a (Laurent) polynomial in $\lambda$.
The coefficients include both positive and negative powers of $\lambda$, which explains why small values of $\lambda$ can lead to significant roundoff error.
For Bini's algorithm all coefficients are monomial with degree between $-1$ and $1$.
